---

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.align = 'center'}
library(ggplot2)
fileLoc <- "C:\\1-vanderbilt\\academic\\seminar\\AirQualityUCI"

raw_data <- read.csv(paste(fileLoc, "AirQualityUCI.csv", sep = '/'), stringsAsFactors = FALSE)
raw_data <- raw_data[1:9357,]
```

# What the data look like

* Dimention and variables

```{r, echo=FALSE, message=FALSE, warning = FALSE, fig.align = 'center'}
raw_data <- raw_data[,-(16:17)]

names(raw_data) <- c('Date','Time','CO_GT',"S1_CO",'NMHC_GT','C6H6_GT','S2_NMHC','NOx_GT','S3_NOx','NO2_GT',"S4_NO2",'S5_O3','T','RH','AH')
droplist <- c('NMHC_GT','C6H6_GT','NOx_GT','NO2_GT')
raw_data <- raw_data[!(names(raw_data)%in%droplist)]
str(raw_data)
kable(head(raw_data))
```

---

# What the data look like

* Missing values analysis

How many missing values for each variable.

```{r, echo = FALSE, message=FALSE, warning = FALSE, fig.align = 'center'}
count_na <- function(list){
  number <- sum(list == -200)
}

number.of.na <- t(sapply(raw_data, count_na))
kable(as.data.frame(number.of.na), format = 'markdown')

raw_data_1 <- raw_data[,1:2]
raw_data_2 <- raw_data[,-(1:2)]

for(val in names(raw_data_2)){
  raw_data_2[val] <- replace(raw_data_2[val], raw_data_2[val] == -200, -1)
}

mod_data <- cbind(raw_data_1, raw_data_2)

```

---

# What the data look like

* Missing values analysis

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE}
mod_data$TS <- paste(mod_data$Date, mod_data$Time, sep = ' ')

mod_data$TS <- strptime(mod_data$TS, "%m/%d/%Y %H:%M:%S", tz = 'UTC')
mod_data$Time <- strptime(mod_data$Time, '%H:%M:%S')
base_line <- strptime('3/9/2004','%m/%d/%Y')
mod_data$Date <- strptime(mod_data$Date,'%m/%d/%Y')
mod_data$day_diff <- round(difftime(mod_data$Date, base_line, units = 'days'))
mod_data$hour <- as.numeric(format(mod_data$Time, '%H'))
```

1. Distribution of missing values of CO_GT

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE}
na_list <- c("CO_GT",'S1_CO','NMHC_GT','NOx_GT','NO2_GT')

data_valid <- mod_data[mod_data['CO_GT'] == -1,]

ggplot(data = data_valid, aes(x = day_diff, y = hour))+
  geom_point(col = 'red') +
  ggtitle("Distributions of missing values across time and date for CO_GT")

ggplot(data = data_valid) +
  geom_bar(mapping = aes(x = hour)) +
  ggtitle("Distributions of missing values across hour for CO_GT")

data_valid <- data_valid[data_valid$hour == '04',]
```

2.Distribution of missing values of S1_CO

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE}
data_valid <- mod_data[mod_data['S1_CO'] == -1,]

ggplot(data = data_valid, aes(x = day_diff, y = hour))+
  geom_point(col = 'red') +
  ggtitle("Distributions of missing values across time and date for multi sensor device")

ggplot(data = data_valid) +
  geom_bar(mapping = aes(x = hour)) +
  ggtitle("Distributions of missing values across hour for multi sensor device")

```

---

# What the data look like

* Missing values analysis

This graph shows the concentration of missing values in detail. for each kind of gas, how many missing values does they each have every day. 

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.cap="", fig.align = 'center'}
library(dplyr)
mod_data$Date <- as.POSIXct(mod_data$Date)
mod_data$Time <- as.POSIXct(mod_data$Time)
mod_data$TS <- as.POSIXct(mod_data$TS)
dataByDay <- mod_data %>%
  group_by(Date) %>%
  summarise(num_CO_GT = sum(CO_GT == -1),
            num_S1_CO = sum(S1_CO == -1))
```

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE,  fig.cap="", fig.align = 'center'}
library(reshape2)
melt_data <- melt(dataByDay, id.vars = 'Date')
ggplot(melt_data, aes(x = Date, y = value, colour = variable)) + geom_point()+ggtitle('distribution of missing values across date')
```

---

# What the data look like

* Missing values analysis

Most missing values are concentrated on several specific days.

Certified reference analyzer did not detect at 4AM everyday. 

They are missing completely at random. 

---

# What the data look like

* How to handle missing values?

Simply delete samples with NA values as we would still have enough samples(9357). 

---

# What the data look like

* Plot time series data of CO_GT(hourly)

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE,  fig.cap="", fig.align = 'center'}
library(lattice)
library(GGally)

for(var in names(mod_data)){
  mod_data[var] = replace(mod_data[var], mod_data[var] == -1, NA)
}

data_without_na <- na.omit(mod_data)

#comp_data <- cbind(mod_data_1, comp_data_2)
y_data <- data_without_na[data_without_na$day_diff %in% c(1,2,3,4,5,6,7),]
ggplot(data = data_without_na, aes(y = CO_GT, x = TS))+
  geom_line()+
  ggtitle('value by hour')

```

To see the detailed pattern.

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.cap="", fig.align = 'center'}
#comp_data <- cbind(mod_data_1, comp_data_2)
y_data <- data_without_na[data_without_na$day_diff %in% c(1,2,3,4,5,6,7,8,9,10),]
ggplot(data = y_data, aes(y = CO_GT, x = TS))+
  geom_line()+
  ggtitle('value by hour within 10 days')

```

---

# What the data look like

* Get the daily average data

Hourly data:

- Too many conseuctive missing values

- Influence of peak hours


```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.cap="", fig.align = 'center'}
comp_data_by_day <- data_without_na %>%
  group_by(Date) %>%
  summarise(day_diff = mean(day_diff),
            co_gt = mean(CO_GT),
            s1_co = mean(S1_CO),
            s2_nmhc = mean(S2_NMHC),
            s3_nox = mean(S3_NOx),
            s4_no2 = mean(S4_NO2),
            s5_o3 = mean(S5_O3),
            t = mean(T),
            rh = mean(RH),
            ah = mean(AH))

ggplot(data = comp_data_by_day, aes(y = co_gt, x = day_diff))+
  geom_line()+
  ggtitle('Daily concentration of CO')

```

