---

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.align = 'center'}
library(ggplot2)
fileLoc <- "C:\\1-vanderbilt\\academic\\seminar\\AirQualityUCI"

raw_data <- read.csv(paste(fileLoc, "AirQualityUCI.csv", sep = '/'), stringsAsFactors = FALSE)
raw_data <- raw_data[1:9357,]

raw_data <- raw_data[,-(16:17)]

names(raw_data) <- c('Date','Time','CO_GT',"S1_CO",'NMHC_GT','C6H6_GT','S2_NMHC','NOx_GT','S3_NOx','NO2_GT',"S4_NO2",'S5_O3','T','RH','AH')
droplist <- c('NMHC_GT','C6H6_GT','NOx_GT','NO2_GT')
raw_data <- raw_data[!(names(raw_data)%in%droplist)]

count_na <- function(list){
  number <- sum(list == -200)
}

number.of.na <- t(sapply(raw_data, count_na))

raw_data_1 <- raw_data[,1:2]
raw_data_2 <- raw_data[,-(1:2)]

for(val in names(raw_data_2)){
  raw_data_2[val] <- replace(raw_data_2[val], raw_data_2[val] == -200, -1)
}

mod_data <- cbind(raw_data_1, raw_data_2)

mod_data$TS <- paste(mod_data$Date, mod_data$Time, sep = ' ')

mod_data$TS <- strptime(mod_data$TS, "%m/%d/%Y %H:%M:%S", tz = 'UTC')
mod_data$Time <- strptime(mod_data$Time, '%H:%M:%S')
base_line <- strptime('3/9/2004','%m/%d/%Y')
mod_data$Date <- strptime(mod_data$Date,'%m/%d/%Y')
mod_data$day_diff <- round(difftime(mod_data$Date, base_line, units = 'days'))
mod_data$hour <- as.numeric(format(mod_data$Time, '%H'))

na_list <- c("CO_GT",'S1_CO','NMHC_GT','NOx_GT','NO2_GT')

data_valid <- mod_data[mod_data['CO_GT'] == -1,]
data_valid <- data_valid[data_valid$hour == '04',]

library(dplyr)
mod_data$Date <- as.POSIXct(mod_data$Date)
mod_data$Time <- as.POSIXct(mod_data$Time)
mod_data$TS <- as.POSIXct(mod_data$TS)
dataByDay <- mod_data %>%
  group_by(Date) %>%
  summarise(num_CO_GT = sum(CO_GT == -1),
            num_S1_CO = sum(S1_CO == -1))

library(reshape2)
#library(mice)
library(lattice)
library(GGally)

for(var in names(mod_data)){
  mod_data[var] = replace(mod_data[var], mod_data[var] == -1, NA)
}

data_without_na <- na.omit(mod_data)

comp_data_by_day <- data_without_na %>%
  group_by(Date) %>%
  summarise(day_diff = mean(day_diff),
            co_gt = mean(CO_GT),
            s1_co = mean(S1_CO),
            s2_nmhc = mean(S2_NMHC),
            s3_nox = mean(S3_NOx),
            s4_no2 = mean(S4_NO2),
            s5_o3 = mean(S5_O3),
            t = mean(T),
            rh = mean(RH),
            ah = mean(AH))
```

# Linear regression

* Correlation and density

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.cap="", fig.align = 'center'}
mod_data_1 <- comp_data_by_day[,c(1,2,12,14,13)]
mod_data_2 <- comp_data_by_day[,-c(1,2,12,14,13)]

my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}

g = ggpairs(na.omit(mod_data_2),lower = list(continuous = my_fn))

g

```

---

# Linear regression

* Transformation

Take log and square root. 

```{r, echo = FALSE, results='hide', message=FALSE, warning = FALSE, fig.cap="", fig.align = 'center'}

plot(density(sqrt(comp_data_by_day$co_gt)), main = 'CO_GT with square root')
plot(density(log(comp_data_by_day$s1_co)), main = 'S1_CO with log')
plot(density(log(comp_data_by_day$s3_nox)),main = 'S3_NOx with log')
plot(density(sqrt(comp_data_by_day$s5_o3)), main = 'S5_O3 with square root')

```

Result was not good.
So no transformation was taken finally.

---

# Linear regression

* adf test

If p-value = 0.01, the actual p-value is smaller than printed one.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tseries)
adf_co <- adf.test(comp_data_by_day$co_gt, k = 3)
adf_s1 <- adf.test(comp_data_by_day$s1_co, k = 3)
adf_s2 <- adf.test(comp_data_by_day$s2_nmhc, k = 3)
adf_s3<- adf.test(comp_data_by_day$s3_nox, k = 3)
adf_s4 <- adf.test(comp_data_by_day$s4_no2, k = 3)
adf_s5 <- adf.test(comp_data_by_day$s5_o3, k = 3)
adf_t <- adf.test(comp_data_by_day$t, k = 3)
adf_rh <- adf.test(comp_data_by_day$rh, k = 3)
adf_ah <- adf.test(comp_data_by_day$ah, k = 3)

adf_list <- c(adf_co$p.value, adf_s1$p.value, adf_s2$p.value, adf_s3$p.value, adf_s4$p.value, adf_s5$p.value, adf_t$p.value, adf_rh$p.value, adf_ah$p.value)

mat1 <- matrix(1:9, nrow = 1, ncol = 9, dimnames=list("p_value",c("co_gt","s1_co","s2_nmhc",                                                                  "s3_nox","s4_no2","s5_o3","t","rh","ah")))
for(i in 1:9){
  mat1[1,i] <- adf_list[i]
}

kable(mat1)
```


---

# Linear regression

* Steps

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(nlme)

data2 <- comp_data_by_day[,c(3,4,5,6,7,8,10,11)]
dataframe_diff <- as.data.frame(sapply(comp_data_by_day, diff))
fit_lm <- lm(co_gt~s1_co+s2_nmhc+s3_nox+s4_no2+s5_o3+rh+ah, data = data2)

step(fit_lm, data = data2, direction = 'backward')

data3 <- data2
data3$co_gt <- sqrt(data3$co_gt)
data3$s1_co <- log(data3$s1_co)
data3$s3_nox <- log(data3$s3_nox)
data3$s5_o3 <- sqrt(data3$s5_o3)

fit_lm2 <- lm(co_gt~s1_co+s2_nmhc+s3_nox+s4_no2+s5_o3+rh+ah, data = data3)

step(fit_lm, data = data3, direction = 'backward')

```

---

# Linear regression

* Regression

```{r, echo = FALSE, message = FALSE, warning = FALSE}
fit_1 <- lm(co_gt ~ s1_co + s2_nmhc + s3_nox + s4_no2 + rh, data = data2)
summary(fit_1)

fit_2 <- lm(co_gt ~ s1_co + s2_nmhc + s3_nox + s4_no2 + rh, data = data3)
summary(fit_2)

```

co_gt has positive relationships with the co, nmhc, nox and ah.

co_gt has negative relationships with no2.

---

# Linear regression

* Estimate and Residuals


```{r, echo = FALSE, message = FALSE, warning = FALSE}
plot.ts(fit_1$fitted.values)

plot.ts(fit_1$residuals)



```

---

# Linear regression

* Estimate and Residuals

```{r, echo = FALSE, message = FALSE, warning = FALSE}
Box.test(fit_1$residuals)

Box.test(fit_2$residuals)

```

---

# Linear regression

* Problem

Residuals are not white noise. 